### 2015-05-27 implementation.tex
Bastien's remarks

\chapter{Implementation}
This chapter will present how we implemented the protocols of chapter~\ref{chap:theory} on the board.
The figure~\ref{fig:os-path-generic} shows the generic data flow in the operating system through the user and kernel spaces.
We will first detail the software part, with the OpenVPN and openSSH as the applications, and OpenSSL as a cryptographic library on which both applications rely.
Then will come strongswan, a user-space abstraction layer giving access to IPsec and we will see how different it is from an OpenVPN implementation.
Lastly, the standard Linux cryptographic kernel modules and network drivers will be listed and briefly discussed.

Before closing the chapter, we will present the two main IPs to which the cryptographic operations will be offloaded and the associated drivers.

@@

Uses regular TCP/UDP network protocols, which can be an advantage over IPsec if the ISP decides to block the latter.

@@

Since OpenVPN aims at network security at the application level, it needs virtual network interface to which to push and from which to read the secured data.
The \texttt{tun} adapter will be used when OpenVPN wants to establish a point-to-point connection between two client, while the \texttt{tap} device is used when there is one server acting as a bridge, hence managing lots of clients at the same time.

OpenVPN as two choices when it wants to sends its data over the internet: either encapsulating its IP packet into TCP or UDP packets.
The problem of encapsulating IP into TCP is that inside IP packets are already TCP frames.
The reason is that IP has been designed to work on unreliable networks, and TCP includes in its standard protocols to retry and drop packets.
Hence, encapsulating an IP packet into a TCP packet produces a redundancy by nesting reliability layers.
UDP is the unreliable counterparty of TCP, offering a better alternative for encapsulation.
If we refer to the figure~\ref{fig:openvpn-packet-flow}, \texttt{iperf} would send its regular \texttt{IP(TCP)} packet over the virtual \texttt{tun} device, which will then compress, fragment and encrypt the frame before encaspulating it into a new UDP packet that will finally be sent through the physical \texttt{eth0} interface.

@@

An other way to look at this workflow is to completely isolate the OpenVPN part from both hosts, and exporting it to an independent device such as a router.
The hosts would be stripped off the virtual interface -- \texttt{tun0} would become physical, and would communicate normally over the internet, at least from their point of view.
All their communications would be routed by the router of their local area network, incomming on a physical \texttt{tun0} interface and outputed to the internet on the \texttt{eth0} interface.
Such implementation are used when an adminstrator wishes to secure a whole local area network without bloating each client with a VPN application.
In this case, there exist router firmware packages running OpenVPN, such as DD-WRT and OpenWRT.
Whatever the method used, the crucial point is that the VPN tunnel usage is transparent to the application.\newline{}

@@

\noindent The first liberty is a laste\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in october 2014.}.
Even if it is not recommended, the user can still enable this mode by explicitly configuring it in the \texttt{sshd\_config} options.

\noindent The second liberty has been taken to prevent the user to strip himself from data authenticity and integrity.
However, in the case of testing and benchmarking, leaving the MAC behind can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
In order to add this feature to OpenSSH, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\ref{chap:openssh-patch}.

@@

Offers an high-level interface called EVP to be used by other applications.

It is to be noted that the present work uses an implementation with all the debug flags activated.
Figure~\ref{fig:openssl-speed-dbg-on-off} shows that if it does have an impact on the performance, it is minimal: in the worst case of the benchmark, the throughput drops only by 2.4\%.
Moreover, the benchmark maximizes this difference by doing only OpenSSL operations.
When OpenSSL will have to share the CPU with other applications, the loss will be even less noticeable.
\begin{figure}[ht]
\begin{tikzpicture}
\begin{axis}[
		title = {Openssl speed with debug on/off},
        width  = 0.85*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar,
        bar width=8pt,
        ymajorgrids = true,
        ylabel = {Throughput [$10^3$kB/s]},
        xlabel = {Packet size},
        ymin=0,
        symbolic x coords={16,64,256,1024,8192},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)},	anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]
\addplot[style={MidnightBlue,fill=MidnightBlue,mark=none}]
	coordinates {(16,18400.60) (64,20750.61) (256,21507.33) (1024,21681.15) (8192,21725.18)};
	\label{aes-256-cbc-dbg-off}

@@

\subsection{Strongswan}
% TODO talk about IKE and SA management
Strongswan is a full implementation of IPsec relying on the kernel drivers for the networking part, on the crypto API for the cryptographic part, and on user space crypto libraries for the connection negociation.
An other popular implementation is ipsec-tools, but its development lags behind modern Linux and is not up-to-date with the 3.14 Linux kernel headers, making its cross-compilation painfull.
%Use strongswan 5.3.0
Strongswan as two advantages: it has a tremendous and exhaustive documentation, and its uer interface is straightforward.
Once configured, a simple \texttt{ipsec start \&\& ipsec up <connection>} on both sides is enough to create a ready to use VPN.

The figure~\ref{fig:ipsec-workflow} illustrates the workflow of Alice communicating with Bob via an IPsec ESP tunnel.
The XFRM, read ``transform", framework is implementing IPsec and handles the incomming and outgoing packets for established VPNs~\cite{rosen2014}.
Its name comes from the fact that the kernel transforms packet frames to incorporate IPsec security.
Depending on the configuration, XFRM uses the AH or ESP kernel module, which in turn calls the crypto API to encrypt and/or sign the IP packet.

@@

The table~\ref{tab:ip-ciphers} summarizes the ciphers supported by the two Barco Silex' IPs used in this work.

\begin{table}[ht]
\begin{tabular}{|l|c|}\hline
IP & Ciphers \\ \hline
BA414E & RSA, DH, DSA, EC \\ 
BA411E & AES modes CBC, CTR, GCM, CCM, CFB, OFB, CTS, ECB \\ \hline
\end{tabular}
\caption{Summary of the ciphers supported by two of Barco Silex' IPs.}{}
\label{tab:ip-ciphers}
\end{table}

\subsection{BA411E Driver}
Takes the rightern path, which is the cleanest because the most versatile.
By pluging the driver into the crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space application via the cryptodev driver and OpenSSL engine.

@@

Be it IRQ or polling, the inner workings are the same and follow the exact same canvas as the default software implementation the different AES modes.

\subsection{BA414E Driver and Silex engine}
At the moment of development, there is no asymetric cryptography interface in the crypto API\footnote{A request for comment patch as been submitted to the Linux kernel mailing list~\cite{crypto-api-pk-encryption} in late April 2015, proposing a standard interface for public key encryption in the crypto API.}.
The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.
As presented in section~\ref{sec:theory-driver}, we will need to rely on an UIO driver to access the device from the user space.

>>>>

\chapter{Implementation}
This chapter will present how we implemented the protocols presented in chapter~\ref{chap:theory} on board.
The figure~\ref{fig:os-path-generic} shows the generic data flow in the operating system through the user and kernel spaces.
We will first detail the software part, with the OpenVPN and openSSH as the applications, and OpenSSL as a cryptographic library on which both applications rely.
Then will come strongswan, a user-space abstraction layer giving access to IPsec and we will see how different it is from an OpenVPN implementation.
Lastly, the standard Linux cryptographic kernel modules and network drivers will be listed and briefly discussed.

Before closing the chapter, we will present the two main IP cores to which the cryptographic operations will be offloaded and the associated drivers.

@@

It uses regular TCP/UDP network protocols, which can be an advantage over IPsec if the ISP decides to block the latter.

@@

Since OpenVPN aims at network security at the application level, it needs virtual network interface to read and write the data.
The \texttt{tun} adapter will be used when OpenVPN wants to establish a point-to-point connection between two client, while the \texttt{tap} device is used when there is one server acting as a bridge, hence managing lots of clients at the same time.

OpenVPN as two choices when it wants to sends its data over the internet: either encapsulating its IP packet into TCP or UDP packets.
The problem of encapsulating IP into TCP is that inside IP packets, there already are TCP frames.
The reason is that IP has been designed to work on unreliable networks, and TCP includes in its standard protocols to retry and drop packets.
Hence, encapsulating an IP packet into a TCP packet produces a redundancy by nesting reliability layers.
UDP is the unreliable counterparty of TCP, offering a better alternative for encapsulation.
If we refer to the figure~\ref{fig:openvpn-packet-flow}, \texttt{iperf} would send its regular \texttt{IP(TCP)} packet over the virtual \texttt{tun} device, which will then compress, fragment and encrypt the frame before encaspulating it into a new UDP packet that will finally be sent through the physical \texttt{eth0} interface.

@@

An other way to look at this workflow is to completely isolate the OpenVPN block from both hosts, and exporting it to an independent device such as a router.
The hosts would be stripped off the virtual interface -- \texttt{tun0} would become physical, and would communicate normally over the internet, at least from their point of view.
All their communications would be routed by the router of their local area network, incomming on a physical \texttt{tun0} interface and outputed to the internet on the \texttt{eth0} interface.
Such implementation are used when an adminstrator wishes to secure a whole local area network without bloating each client with a VPN application.
In this case, there exist router firmware packages running OpenVPN, such as DD-WRT and OpenWRT.
Whatever the method used, the crucial point is that the VPN tunnel usage is transparent to the application.\newline{}

@@

\noindent The first customization is a response to an old vulnarability\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in october 2014.}.
Even if it is not recommended, the user can still enable this mode by explicitly configuring it in the \texttt{sshd\_config} options.

\noindent The second has been taken to prevent the user to strip himself from data authenticity and integrity.
However, in the case of testing and benchmarking, skipping the MAC can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
In order to overcome this limitation, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\ref{chap:openssh-patch}.

@@

OpenSSL offers an high-level interface called EVP to be used by other applications.

It is to be noted that the present work uses an implementation with all the debug flags activated.
Figure~\ref{fig:openssl-speed-dbg-on-off} shows that if it does have an impact on the performance, it is negligible for our tests: in the worst case of the benchmark, the throughput drops only by 2.4\%.
Moreover, the benchmark maximizes this difference by doing only OpenSSL operations.
When OpenSSL will have to share the CPU with other applications, the loss will be even less noticeable.
\begin{figure}[ht]
\begin{tikzpicture}
\begin{axis}[
		title = {Openssl speed with debug on/off},
        width  = 0.85*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar,
        bar width=8pt,
        ymajorgrids = true,
        ylabel = {Throughput [$10^3$kB/s]},
        xlabel = {Packet size},
        ymin=0,
        symbolic x coords={16,64,256,1024,8192},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)},	anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]
\addplot[style={MidnightBlue,fill=MidnightBlue,mark=none}]
	coordinates {(16,18400.60) (64,20750.61) (256,21507.33) (1024,21681.15) (8192,21725.18)};
	\label{aes-256-cbc-dbg-off}

@@

\subsection{Strongswan}
% TODO talk about IKE and SA management
Strongswan is a full implementation of IPsec relying on the kernel drivers for the networking part, on the Linux Crypto API for the cryptographic part, and on user space crypto libraries for the connection negociation.
An other popular implementation is ipsec-tools, but its development lags behind modern Linux and is not up-to-date with the 3.14 Linux kernel headers, making its cross-compilation difficult.
%Use strongswan 5.3.0
Strongswan as two advantages: it has a tremendous and exhaustive documentation, and its user interface is straightforward.
Once configured, a simple \texttt{ipsec start \&\& ipsec up <connection>} on both sides is enough to create a ready-to-use VPN.

The figure~\ref{fig:ipsec-workflow} illustrates the workflow of Alice communicating with Bob via an IPsec ESP tunnel.
The XFRM (read ``transform") framework is implementing IPsec and handles the incomming and outgoing packets for established VPN connections~\cite{rosen2014}.
Its name comes from the fact that the kernel transforms packet frames to incorporate IPsec security.
Depending on the configuration, XFRM uses the AH or ESP kernel module, which in turn calls the crypto API to encrypt and/or sign the IP packet.

@@

The table~\ref{tab:ip-ciphers} summarizes the ciphers supported by the two Barco Silex' IP cores used in this work.

\begin{table}[ht]
\begin{tabular}{|l|c|}\hline
IP core & Ciphers \\ \hline
BA414E & RSA, DH, DSA, ECDSA, ECDH \\ 
BA411E & AES modes CBC, CTR, GCM, CCM, CFB, OFB, XTS, ECB \\ \hline
\end{tabular}
\caption{Summary of the ciphers supported by two of Barco Silex' IP cores.}{}
\label{tab:ip-ciphers}
\end{table}

For this tests, only one instance of each IP core will be programmed on the FPGA.
As all the tests will be single-threaded, having two or more instances of the same IP core would not improve the performance, as they could not be used in parallel anyway.

\subsection{BA411E Driver}
This driver takes the rightern path of figure~\ref{fig:os-path-generic-b}, which is the best because the most versatile.
By pluging the driver into the Linux Crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space application via the cryptodev driver and OpenSSL engine.

@@

Whichever method is used to fetch the state of the hardware, the rest of the driver follow the exact same canvas as the default software implementation of the different AES modes.

\subsection{BA414E Driver and Silex engine}
At the moment of development, there is no asymmetric cryptography interface in the Linux Crypto API\footnote{A request for comment patch as been submitted to the Linux kernel mailing list~\cite{crypto-api-pk-encryption} in late April 2015, proposing a standard interface for public key encryption in the crypto API.}.
The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.
As presented in section~\ref{sec:theory-driver}, we will need to rely on an UIO driver to access the device from the user space.

### 2015-05-27 technical-background.tex
Bastien's remarks

This chapter will address the technical ground inherent to this work.
First comes an overview of Linux operating systems, the distinction between user and kernel mode, and the design of device drivers.

\noindent Follow a quick presentation of FPGAs and how they can be driven from the operating system.

\noindent Next comes An overview of the cryptographic field with the main primitives: message digests, symmetric ciphers and asymmetric ciphers.

\noindent Finally, the ground necessary to understand the inner working and challenges of some VPNs implementations will be covered with SSL/TLS and IPsec.

@@

There are two ways to get the results from a hardware device: either by using interruption, or by actively polling the device, relentlessly asking it if it finished its operations.
The first case is the cleanest and the most common: when the device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will in turn execute the interruption routine registered by the driver~\citep[chap. 10]{Corbet:2005:LDD:1209083}.
The second case is the easiest and is always guaranteed to work, but won't let go off the processor willingly, loading it at 100\%, and avoiding any other task to be executed.
Hopefully, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, that is the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.

@@

Cryptography is the corner stone of security.
The four main goals are the following, as defined in~\cite{Menezes1996}:
\begin{description}
	\item[Confidentiality] keeping information secret from all but those who are authorized to see it.
	\item[Integrity] ensuring information has not been altered by unauthorized or unknown means.
	\item[Source Authentication] corroborating the source of information.
	\item[Non-repudiation] preventing the denial of previous commitments or actions.
\end{description}

In order to achieve those, four cryptographic primitives are needed: symmetric and asymmetric ciphers, message digests and digital signatures.

\subsection{Message digest}
%Not the same as digital signature because we use the same key for both MAC values.
A message digest is the result of a one-way mathematical function of a fixed size.
Those hash functions are of two types~\cite{infof405}: manipulation detection codes (MDC) to guarantee integrity and message authentication codes (MAC) to guarantee both integrity and source authentication.

@@

Based on this design and adding a key to the process, the RFC 2104~\cite{rfc2104} defines a MAC:
\[
	HMAC(k, x) = h((k\oplus opad)|h((k\oplus ipad)|x))
\]
with a key $k$, and two padding block added for security concerns: an outer pad $opad$ and an inner pad $ipad$.

@@

\subsection{Symetric cryptography}
Talk about encryption, integrity and authentication.

@@

\subsection{Asymetric cryptography}
% Diffie-Hellman: show the math about the shared key, maybe from the course INFOH405 or RFC2631, and point which operation will be offloaded.
Asymetric cryptography relies on a pair of keys: one private known only to the owner of the certificate, and one public available to anyone.
Such cryptography uses two kinds of operations: encryption using the public key of the recipient and digital signature, which is an ecnryption using the private key of the sender.

\subsubsection{RSA}
RSA is a public-key scheme proposed in 1978 by three MIT researchers who gave it their name~\cite{Rivest:1978:MOD:359340.359342}.
A few years later, they founded RSA Laboratories, which is now in charge of maintaining its standards, alongside many others, as the first Public-Key Cryptography Standards, \textit{aka} PKCS \#1.
The last version of the standard is the version 2.2~\cite{pkcs1} and is defined as a precise key generation protocol allowing encryption and decryption.
The keys can be generated by respecting a few steps:
\begin{enumerate}
	\item randomy choose two large primes $p$ and $q$;
	\item compute the modulus $n = p q$, and consequently we have $\phi(n) = (p-1)(q-1)$, with $\phi(n)$ as the Euler function;
	\item randomly choose the public exponent $e \in ]1,\phi(n)[\ s.t.\ GCD(e,\phi(n)) = 1$;
	\item compute $d \in ]1,\phi(n)[\ s.t.\ e \cdot d \equiv 1 (mod\ \phi(n))$
\end{enumerate}

@@

\subsubsection{Diffie-Hellman}
Diffie-Hellman is a secret key exchange protocol: two parties compute a shared secret $ZZ$ that can be used as a symetric key during the following exchanges.
It uses the same kind of operation as RSA, that is modular exponentiation.
The protocol can be one of two type (\cite{rfc2631}, \cite{Frankel:2005:SGI:2206289}):
\begin{itemize}
	\item Static: the actors use their authenticated certificate to compute the shared secret.
	\item Ephemeral: the actors create a new pair of public/private keys from which the secret key is derived.
	\item Anonymous: same as ephemeral, but without signing anything, hence not identifying neither of the actors. This mode is not advisable since it's vulnerable to main-in-the-middle attack.
\end{itemize}
A static scheme is easier to implement and requires much less operations, but using ephemeral keys is essential to ensure perfect forward secrecy.
Imagine that somehow, an opponent lays his hand on the shared secret.
If that secret has already been used, he can decipher all data transfered during past connections.
However, if the secret is new for every new connection, the compromission of the shared secret des not jeopardize past communications.
This is perfect forward secrecy: using a new key to protect the past.

Hereunder is the generation of an ephemeral shared secret.
For a static secret, Alice and Bob will simply use their static certificate, sparing the modular exponentiation of the ephemeral public key generation.
\begin{enumerate}
	\item Alice generates once $p$ and $g$ (using precomputed parameters):
	\begin{description}[nosep]
		\item[p] large prime number
		\item[g] a generator of $\mathds{Z}_p^*$
	\end{description}
	\item Alice picks a random integer $x_a$ and computes $g^{x_a} mod\ p = y_a$.
	\item Alice sends $p$, $g$ and $y_a$ to Bob, signing everything using her private certificate.
	\item Bob checks the signature and picks $x_b$.
	\item Bob computes $y_a^{x_b} mod\ p = g^{x_a x_b} mod\ p = ZZ$, the shared secret to use as a premaster key from which will be derived the symetric key for further communications.
	\item Bob sends $y_b = g^{x_b} mod\ p$, signing everything with his private certificate.
	\item Alice checks the signature and computes the same shared secret: $ZZ = y_b^{x_a} mod\ p = g^{x_a x_b} mod\ p$
\end{enumerate}

If the server is Alice, it has to do at least one signature, one signature verification and two modular exponentiations.
Note that the client, B in our case, could have one signature less because RFC 5246~\cite{rfc5246} leave it as an optional feature, and the server would then have one verification less.
However, any sane configuration will have both actors signing their ephemeral public key.
If the certificate use RSA, we end up with four modular exponentiations, which can become quite heavy computing wise for certain sizes of prime numbers.
We will see in chapter~\ref{chap:results} that while a 1024-bit prime is easily manageable by full software implementation, hardware offloading become a necessity for 4096-bit primes.
Moreover, 1024-bit parameter size, both RSA and Diffie-Hellman, are disallowed by the NIST recommandations since 2013~\cite{nist-sp800-131A}.

@@

\begin{table}[ht]
\center
\begin{tabularx}{\textwidth}{|l|l|l|X|} \hline
\multicolumn{2}{|c|}{TCP/IP layering} & \multicolumn{2}{c|}{OSI model} \\ \hline
Layer & Protocols & Layer & Protocols \\ \hline
\multirow{3}{*}{Application} & \multirow{3}{*}{FTP, SSH} & Application & FTP \\ \cline{3-4}
 & & Presentation & ASCII, JPEG \\ \cline{3-4}
 & & Session & RPC, PAP \\ \hline
Transport & TCP, UDP & Transport & TCP, UDP \\ \hline
Internet & IP, ICMP, IPsec & Network & IP, ICMP, IPsec \\ \hline
\multirow{2}{*}{Data Link} & \multirow{2}{*}{PPP, MAC, Ethernet, L2TP} & Data link & PPP, MAC, Ethernet, L2TP \\ \cline{3-4}
 & & Physical & USB, DSL, IEEE 802.11 \\ \hline
\end{tabularx}
\caption{TCP/IP and OSI model comparison}{They are globaly the same, except for the application layer of the TCP/IP stack which merge togheter the three upper layers of the OSI model. Between parenthsis are examples of protocols resting on each layer.}
\label{tab:tcp-ip-stack}
\end{table}

@@

@@

\subsection{IPsec}
% IPsec as a protocol, strongswan come in the 'implementation' chapter.
IPsec is a modification of the IP stack in the kernel space.
It is a network/internet level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.

>>>>

This chapter will address the technical ground necessary to this work.
First comes a quick presentation of FPGAs and how they can be driven from the operating system.

\noindent Follows an overview of Linux operating systems, the distinction between user and kernel mode, and the design of device drivers.

\noindent Next comes an overview of the cryptographic field with the main primitives: message digests, symmetric ciphers and asymmetric ciphers.

\noindent Finally, the ground necessary to understand the inner working and challenges of some VPNs implementations using SSL/TLS and IPsec will be covered.

+ talk about the slave/master and the two data paths: data and control.

@@

There are two ways to get the status of a hardware device: either by using interruption, or by actively polling the device, relentlessly asking if it finished its operations.
The first case is the cleanest and the most common: when the device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will in turn execute the interruption routine registered by the driver~\citep[chap. 10]{Corbet:2005:LDD:1209083}.
The second case is the easiest and is always guaranteed to work, but won't let go off the processor willingly, loading it at 100\%, and avoiding any other task to be executed.
Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, that is the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.

@@

Cryptography is the keystone of security. As \citet{Menezes1996} defines it: ``Cryptography is the study of mathematical techniques related to aspects of information security such as confidentiality, data integrity, entity authentication, and data origin authentication."

The four main goals are:
\begin{description}
	\item[Confidentiality] keeping information secret from all but those who are authorized to see it.
	\item[Integrity] ensuring information has not been altered by unauthorized or unknown means.
	\item[Source Authentication] corroborating the source of information.
	\item[Non-repudiation] preventing the denial of previous commitments or actions.
\end{description}

In order to achieve those, four cryptographic primitives are needed: message digests, symmetric and asymmetric ciphers, and digital signatures.

\subsection{Message digest}
%Not the same as digital signature because we use the same key for both MAC values.
A message digest is the result of a one-way mathematical function of a fixed size.
For a given function, each message has a unique digest, comparable to a fingerprint.
Those hash functions are of two types~\cite{infof405}: manipulation detection codes (MDC) to guarantee integrity and message authentication codes (MAC) to guarantee both integrity and source authentication.

@@

In order to create a MAC, we can add a secret key $k$ to the process, $H_i$ becoming:
\[
	H_i = f(H_{i-1}, x_i, k), \mbox{with } i \in [1,t]\\
\]

HMAC is such function $f()$, as defined in RFC 2104~\cite{rfc2104}
\[
	HMAC(k, x) = h((k\oplus opad)|h((k\oplus ipad)|x))
\]
with a key $k$, and two padding block added for security concerns: an outer pad $opad$ and an inner pad $ipad$.

@@

\subsection{Symmetric cryptography}
Talk about encryption, integrity and authentication.

@@

\subsection{Asymmetric cryptography}
% Diffie-Hellman: show the math about the shared key, maybe from the course INFOH405 or RFC2631, and point which operation will be offloaded.
Asymmetric cryptography relies on a pair of keys: one private known only to the owner of the certificate, and one public available to anyone.
Such cryptography uses two kinds of operations: encryption using the public key of the recipient and digital signature, which is an encryption using the private key of the sender.

\subsubsection{RSA}
RSA is a public-key scheme proposed in 1978 by three MIT researchers (who gave it their name)~\cite{Rivest:1978:MOD:359340.359342}.
A few years later, they founded RSA Laboratories, which is now in charge of maintaining its standards, alongside many others, as the first Public-Key Cryptography Standards, \textit{aka} PKCS \#1.
The last version of the standard is the version 2.2~\cite{pkcs1} and is defined as a precise key generation protocol allowing encryption and decryption.
The keys can be generated by respecting a few steps:
\begin{enumerate}
	\item randomy choose two large primes $p$ and $q$;
	\item compute the modulus $n = p q$, and consequently we have \[\phi(n) = (p-1)(q-1)\]with $\phi(n)$ as the Euler function;
	\item randomly choose the public exponent $e \in ]1,\phi(n)[\ s.t.\ GCD(e,\phi(n)) = 1$;
	\item compute $d \in ]1,\phi(n)[\ s.t.\ e \cdot d \equiv 1 (mod\ \phi(n))$
\end{enumerate}

@@

The scheme is based on modular artihmetic, which is secure because of the factoring problem.
For an opponent to decrypt a ciphertext $c$ wihtout knowing the private key $(d, n)$, he would need to compute the $e^{th}$ roots of $c$, modulo $n$.
This operation is computationally hard, since an efficient solution is yet to be found for sufficiently large key size.

\subsubsection{Diffie-Hellman}
Diffie-Hellman is a secret key exchange protocol: two parties compute a shared secret $ZZ$ that can be used as a symmetric key during the following exchanges.
It uses the same kind of operation as RSA, that is modular exponentiation.
The protocol can be one of three types (\cite{rfc2631}, \cite{Frankel:2005:SGI:2206289}):
\begin{itemize}
	\item Static (DH): the actors use their authenticated certificate to compute the shared secret.
	\item Ephemeral (EDH): the actors create a new pair of public/private keys from which the secret key is derived.
	\item Anonymous (ADH): same as ephemeral, but without signing anything, hence not identifying neither of the actors, hence not providing any authentication. This mode is not advisable since it's vulnerable to main-in-the-middle attack.
\end{itemize}
A static scheme is easier to implement and requires much less operations, but using ephemeral keys is essential to ensure perfect forward secrecy.
Imagine that somehow, an opponent lays his hand on the shared secret.
If that secret has already been used, he can decipher all data transfered during past connections.
However, if the secret is new for every new connection, the compromission of the shared secret does not jeopardize past communications.
This is perfect forward secrecy: using a new key to protect the previous messages.

Hereunder is the generation algorithm of an ephemeral shared secret.
For a static secret, Alice and Bob will simply use their static certificate, sparing the modular exponentiation of the ephemeral public key generation.
\begin{enumerate}
	\item Alice generates once $p$ and $g$ (using precomputed parameters):
	\begin{description}[nosep]
		\item[p] large prime number
		\item[g] a generator of $\mathds{Z}_p^*$
	\end{description}
	\item Alice picks a random integer $x_a$ and computes $g^{x_a} mod\ p = y_a$.
	\item Alice sends $p$, $g$ and $y_a$ to Bob, signing everything using her private certificate.
	\item Bob checks the signature and picks $x_b$.
	\item Bob computes $y_a^{x_b} mod\ p = g^{x_a x_b} mod\ p = ZZ$, the shared secret to use as a premaster key from which will be derived the symmetric key for further communications.
	\item Bob sends $y_b = g^{x_b} mod\ p$, signing everything with his private certificate.
	\item Alice checks the signature and computes the same shared secret: $ZZ = y_b^{x_a} mod\ p = g^{x_a x_b} mod\ p$
\end{enumerate}

If the server is Alice, it has to do at least one signature generation, one signature verification and two modular exponentiations.
Note that the client, B in our case, could have one signature less because RFC 5246~\cite{rfc5246} leave it as an optional feature, and the server would then have one verification less.
However, any sane configuration will have both actors signing their ephemeral public key.
If the certificate use RSA, we end up with four modular exponentiations, which can become quite heavy computing wise for certain sizes of key.
We will see in chapter~\ref{chap:results} that while a 1024-bit key size is easily manageable by full software implementation, hardware offloading become a necessity for 4096-bit key sizes.
Moreover, 1024-bit parameter size, both RSA and Diffie-Hellman, are disallowed by the NIST recommandations since 2013~\cite{nist-sp800-131A}.

@@

\begin{table}[ht]
\center
\begin{tabularx}{\textwidth}{|l|l|l|X|} \hline
\multicolumn{2}{|c|}{TCP/IP layering} & \multicolumn{2}{c|}{OSI model} \\ \hline
Layer & Protocols & Layer & Protocols \\ \hline
\multirow{3}{*}{Application} & \multirow{3}{*}{FTP, SSH} & Application & FTP \\ \cline{3-4}
 & & Presentation & ASCII, JPEG \\ \cline{3-4}
 & & Session & RPC, PAP \\ \hline
Transport & TCP, UDP & Transport & TCP, UDP \\ \hline
Internet & IP, ICMP, IPsec & Network & IP, ICMP, IPsec \\ \hline
\multirow{2}{*}{Link} & \multirow{2}{*}{PPP, MAC, Ethernet, L2TP} & Data link & PPP, MAC, Ethernet, L2TP \\ \cline{3-4}
 & & Physical & USB, DSL, IEEE 802.11 \\ \hline
\end{tabularx}
\caption{TCP/IP and OSI model comparison}{They are globaly the same, except for the application layer of the TCP/IP stack which merge togheter the three upper layers of the OSI model. Between parenthsis are examples of protocols resting on each layer.}
\label{tab:tcp-ip-stack}
\end{table}

@@

\noindent L2TP (Layer 2 Tunneling Protocol) is a protocol that simply offers a tunnel without providing confidentiality~\cite{rfc3931}.
A common use case is to combine IPsec and L2TP, as defined in RFC 3193~\cite{rfc3193}.

@@

\subsection{IPsec}
% IPsec as a protocol, strongswan come in the 'implementation' chapter.
IPsec is a protocol whose implementation is a modification of the IP stack in the kernel space.
It is a network/internet level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.
